### 2004-06-28 doc/Simulation_and_Control_using_Supremica/functionblocks.tex
*** empty log message ***

New technologies and standards are emerging that are set to have a
dramatic effect on the design and implementation of industrial
control systems. A new standard IEC-61131-3~\cite{IEC-61131-3} was
introduced in 1995 to standardize the representation of control
designs so as to enable cross-vendor integration. The standard
offered significant advantages over earlier vendor proprietary
control design representations~\cite{Book-61131}.
\begin{itemize}
\item It encouraged well-structured procedural
programs that allowed the development of different functional
elements into separate procedures or program units.
\item It provided standardized languages and methods of program execution
so that wide range of technological problems can be programmed as
vendor-independent software elements.
\item It provided support for various languages like structured
text, function block diagrams, ladder diagrams, instruction list
and sequential function charts.
\item The standard enforced strong data typing and supported user defined
data structures.
\end{itemize}

@@

To overcome these problems, a function block model was sought that
would support distributed control systems in a scalable,
extensible and flexible manner. IEC-61499~\cite{IEC-61499}
standard has now been developed for distributed industrial
process, measurement and control systems.

@@

The following sections are adapted largely from the various
discussions on the features of the standard in ~\cite{IEC-61499,
IEC-61131-3, Book-61131, Book-61499, christensen1, christensen2}

@@

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/System_Model.eps, scale=.8} \caption[System
Model]{System Model {\protect ~\cite{IEC-61499}}}
\label{f:System_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Device_Model.eps, scale=.8} \caption[Device
Model]{Device Model{\protect ~\cite{IEC-61499}}}
\label{f:Device_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Resource_Model.eps, scale=.8} \caption[Resource
Model]{Resource Model{\protect ~\cite{IEC-61499}}}
\label{f:Resource_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Application_Model.eps, scale=.8}
\caption[Application Model]{Application Model{\protect
~\cite{IEC-61499}}} \label{f:Application_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Function_Block_Char.eps, scale=.8}
\caption[Function Block Characteristics]{Function Block
Characteristics{\protect ~\cite{IEC-61499}}}
\label{f:Function_Block_Char}
\end{center}
\end{figure}
%%\doublespace
%

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Basic_Function_Block.eps, scale=.8}
\caption[Basic Function Block]{Basic Function Block{\protect
~\cite{IEC-61499}}} \label{f:Basic_Block}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Composite_Function_Block.eps, scale=.8}
\caption[Composite Function Block]{Composite Function
Block{\protect ~\cite{IEC-61499}}} \label{f:Composite_Block}
\end{center}
\end{figure}
%\doublespace
%

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Execution_Control_Chart.eps, scale=.8}
\caption[Execution Model]{Execution Model{\protect
~\cite{IEC-61499}}} \label{f:Execution_Model}
\end{center}
\end{figure}
%\doublespace
%

@@

\section{FBEditor}
Function Block Development Kit (FBDK) ~\cite{FBEditor} is a
software toolkit that enables users to build and test data types,
function block types, resource types, device types and system
configurations according to the IEC 61499 standard. FBEditor is a
part of the FBDK toolkit that provides a development and runtime
environment for applications based on the IEC 61499 standard.
Figure \ref{f:FBEditor} shows the editor being used to load a
function block used for controlling a motor.

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/FBEditor.eps, scale=.6}
\caption[FBEditor]{FBEditor{\protect ~\cite{FBEditor}}}
\label{f:FBEditor}
\end{center}
\end{figure}
%\doublespace
%

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Devices.eps, scale=.6} \caption[FBEditor
- Editing system devices]{FBEditor - Editing system
devices{\protect ~\cite{FBEditor}}} \label{f:Editor-Devices}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Mapping.eps, scale=.6} \caption[FBEditor
- Editing device connections]{FBEditor - Editing device
connections{\protect ~\cite{FBEditor}}} \label{f:Editor-Mapping}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Versions.eps, scale=.6}
\caption[FBEditor - Editing meta properties]{FBEditor - Editing
meta properties{\protect ~\cite{FBEditor}}}
\label{f:Editor-Versions}
\end{center}
\end{figure}
%\doublespace
%
FBEditor has a navigation pane that allows the user to choose the
display on the worksheet area, including the external interface;
the Execution Control Chart (ECC) for basic function block type;
service primitive sequences for service interface function block
types; and function block networks for composite function block
types, device and resource types and instances, and system
configurations. The worksheet / viewport area contains the
function block diagram for the currently selected instance.
Different components of the diagram can be dragged around for
better readability.

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Transitions.eps, scale=.6}
\caption[FBEditor Editing the execution control chart]{FBEditor
Editing the execution control chart{\protect ~\cite{FBEditor}}}
\label{f:Editor-Transitions}
\end{center}
\end{figure}
%\doublespace
%

>>>>

New technologies and standards are emerging that are set to have a
dramatic effect on the design and implementation of industrial
control systems. A new standard IEC-61131-3~\cite{iec:1131:1993} was
introduced in 1995 to standardize the representation of control
designs so as to enable cross-vendor integration. The standard
offered significant advantages over earlier vendor proprietary
control design representations~\cite{l:pro:1995}.
\begin{itemize}
\item It encouraged well-structured procedural
programs that allowed the development of different functional
elements into separate procedures or program units.
\item It provided standardized languages and methods of program execution
so that wide range of technological problems can be programmed as
vendor-independent software elements.
\item It provided support for various languages like structured
text, function block diagrams, ladder diagrams, instruction list
and sequential function charts.
\item The standard enforced strong data typing and supported user defined
data structures.
\end{itemize}

@@

To overcome these problems, a function block model was sought that
would support distributed control systems in a scalable,
extensible and flexible manner. IEC-61499~\cite{iec:61499:2000}
standard has now been developed for distributed industrial
process, measurement and control systems.

@@

The following sections are adapted largely from the various
discussions on the features of the standard in ~\cite{iec:61499:2000,
iec:1131:1993, l:pro:1995, l:mod:2001, c:des:2002, c:ope:2002}

@@

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/System_Model.eps, scale=.8} \caption[System
Model]{System Model {\protect ~\cite{iec:61499:2000}}}
\label{f:System_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Device_Model.eps, scale=.8} \caption[Device
Model]{Device Model{\protect ~\cite{iec:61499:2000}}}
\label{f:Device_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Resource_Model.eps, scale=.8} \caption[Resource
Model]{Resource Model{\protect ~\cite{iec:61499:2000}}}
\label{f:Resource_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Application_Model.eps, scale=.8}
\caption[Application Model]{Application Model{\protect
~\cite{iec:61499:2000}}} \label{f:Application_Model}
\end{center}
\end{figure}
%%\doublespace
%

%
%%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Function_Block_Char.eps, scale=.8}
\caption[Function Block Characteristics]{Function Block
Characteristics{\protect ~\cite{iec:61499:2000}}}
\label{f:Function_Block_Char}
\end{center}
\end{figure}
%%\doublespace
%

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Basic_Function_Block.eps, scale=.8}
\caption[Basic Function Block]{Basic Function Block{\protect
~\cite{iec:61499:2000}}} \label{f:Basic_Block}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Composite_Function_Block.eps, scale=.8}
\caption[Composite Function Block]{Composite Function
Block{\protect ~\cite{iec:61499:2000}}} \label{f:Composite_Block}
\end{center}
\end{figure}
%\doublespace
%

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Execution_Control_Chart.eps, scale=.8}
\caption[Execution Model]{Execution Model{\protect
~\cite{iec:61499:2000}}} \label{f:Execution_Model}
\end{center}
\end{figure}
%\doublespace
%

@@

\section{FBEditor}
Function Block Development Kit (FBDK) ~\cite{c:fun:2002} is a
software toolkit that enables users to build and test data types,
function block types, resource types, device types and system
configurations according to the IEC 61499 standard. FBEditor is a
part of the FBDK toolkit that provides a development and runtime
environment for applications based on the IEC 61499 standard.
Figure \ref{f:FBEditor} shows the editor being used to load a
function block used for controlling a motor.

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/FBEditor.eps, scale=.6}
\caption[FBEditor]{FBEditor{\protect ~\cite{c:fun:2002}}}
\label{f:FBEditor}
\end{center}
\end{figure}
%\doublespace
%

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Devices.eps, scale=.6} \caption[FBEditor
- Editing system devices]{FBEditor - Editing system
devices{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Devices}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Mapping.eps, scale=.6} \caption[FBEditor
- Editing device connections]{FBEditor - Editing device
connections{\protect ~\cite{c:fun:2002}}} \label{f:Editor-Mapping}
\end{center}
\end{figure}
%\doublespace
%

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Versions.eps, scale=.6}
\caption[FBEditor - Editing meta properties]{FBEditor - Editing
meta properties{\protect ~\cite{c:fun:2002}}}
\label{f:Editor-Versions}
\end{center}
\end{figure}
%\doublespace
%
FBEditor has a navigation pane that allows the user to choose the
display on the worksheet area, including the external interface;
the Execution Control Chart (ECC) for basic function block type;
service primitive sequences for service interface function block
types; and function block networks for composite function block
types, device and resource types and instances, and system
configurations. The worksheet / viewport area contains the
function block diagram for the currently selected instance.
Different components of the diagram can be dragged around for
better readability.

@@

%
%\singlespace
\begin{figure}
\begin{center}
\epsfig{file=images/functionblocks/Editor-Transitions.eps, scale=.6}
\caption[FBEditor Editing the execution control chart]{FBEditor
Editing the execution control chart{\protect ~\cite{c:fun:2002}}}
\label{f:Editor-Transitions}
\end{center}
\end{figure}
%\doublespace
%
